### 2018-08-04 modelling.tex
add network figure
Fix definition of p-value

The results of this analysis will depend not only of the observed data, but also the data generation process \cite{Kruschke}. This way, both tools return point estimates of the \ac{log2FC} between a pair of conditions, and a \textit{p-value}, which measures the probability of the null hypothesis. The null hypothesis usually states that the true value of the parameter is 0. Thus, \textit{p-values} do not say anything about the probability of hypotheses outside $H_0$ and $H_1$. Is there a framework that reports results for a particular hypothesis, based just on the seen data and our prior knowledge?

@@

It can be applied to the quantification problem, where $\theta$ turns into the \ac{log2FC} parameter we try to estimate. All is needed is the computation of the posterior probability distribution. Even though, this is tractable analytically for simple models only, fortunately Markov Chain-Monte Carlo (\ac{MCMC}) and Variational Inference (\ac{VI}) methods can be used to approximate this posterior when the analytical derivation is not possible. Both \ac{MCMC} and \ac{VI} are now possible to run thanks to the advent of modern computing.

@@

On the other hand, \ac{VI} methods, very recently developed and introduced in PyMC3, provide a fast alternative to \ac{MCMC} methods, yet they are not guaranteed to asymptotically approximate the true posterior. In \ac{VI}, optimization, instead of sampling, is employed as the way to approximate the posterior \cite{Blei2017}. More concretely, a family of densities $Q$ over the parameters $\theta$ is defined. The goal of \ac{VI} is to find the single density $q$ within the family $Q$ that best approximates $p(\theta|y)$. This approximate density should be complex enough to reach a good approximation, and at the same time simple enough to be computationally easy to work with. The best candidate is defined as the one minimising the Kullback-Leibler (\ac{KL}) divergence (see equation \ref{eq:KL}).

@@

Both because of the little knowledge on the value of the parameters governing the data generation process and to provide a prior that everyone can agree upon, non informative priors were provided to the model. This was condensed in the specification of Normal and Half Normal distributions for all the random stochastic variables in the model. The hyperparameters were selected to be non-informative.

@@

%\begin{figure}[H]
% \centering 
% \includegraphics{hierarch_diagram.pdf}
% \caption{}
% label{fig:dbda}
%\end{figure}
\begin{figure}[!h]
\input{hierarch_diagram}
\caption[BayesQuant Kruschke-like diagram]{\textbf{BayesQuant Kruschke-like diagram}. Extends figure \ref{fig:daft_model} to give more detail into the model. Prior probability distributions are indicated with "distrograms", and the dependencies between them via arrows. The hyperparameters are represented by latin capital letters. Each $\beta$ will consist of a vector of dimensions $1 \times n$ where $n$ is the number of different factors in the category (i.e 2 for treatment, 6 for run, etc). The corresponding $x$ variable is a matrix of dimensions equal to $n \times n_peptides$ where the $i,j$ cell indicates if the $jth$ peptide is affected by the $ith$ factor (1) or not (0).}
\label{fig:dbda}
\end{figure}

@@

\subsection{Sampling from the approximation to the posterior}
\label{subsec:basic_model}

>>>>

The results of this analysis will depend not only of the observed data, but also the data generation process \cite{Kruschke}. This way, both tools return point estimates of the \ac{log2FC} between a pair of conditions, and a \textit{p-value}, which measures the probability of the data under the null hypothesis. The null hypothesis usually states that the true value of the parameter is 0. Thus, \textit{p-values} do not say anything about the probability of the hypotheses. Is there a framework that reports hypotheses probabilities, based just on the seen data and our prior knowledge?

@@

It can be applied to the quantification problem, where $\theta$ turns into the \ac{log2FC} parameter we try to estimate. All is needed is the computation of the posterior probability distribution. Even though, this is tractable analytically for simple models only, fortunately Markov Chain-Monte Carlo (\ac{MCMC}) and Variational Inference (\ac{VI}) methods can be used to approximate this posterior when the analytical derivation is not possible.

@@

On the other hand, \ac{VI} methods, very recently developed and introduced in PyMC3, provide a fast alternative to \ac{MCMC}, yet they are not guaranteed to asymptotically approximate the true posterior. In \ac{VI}, optimization, instead of sampling, is employed as the way to approximate the posterior \cite{Blei2017}. More concretely, a family of densities $Q$ over the parameters $\theta$ is defined. The goal of \ac{VI} is to find the single density $q$ within the family $Q$ that best approximates $p(\theta|y)$. This approximate density should be complex enough to reach a good approximation, and at the same time simple enough to be computationally easy to work with. The best candidate is defined as the one minimising the Kullback-Leibler (\ac{KL}) divergence (see equation \ref{eq:KL}).

@@

Both because of the little knowledge on the value of the parameters governing the data generation process and to provide a prior that everyone can agree upon, weakly informative priors were provided to the model. This was condensed in the specification of Normal and Half Normal distributions for all the random stochastic variables in the model. The hyperparameters were selected to be non-informative.

@@

%\begin{figure}[H]
% \centering 
% \includegraphics{hierarch_diagram.pdf}
% \caption{}
% label{fig:dbda}
%\end{figure}
\begin{figure}[!h]
\input{hierarch_diagram}
\caption[BayesQuant Kruschke-like diagram]{\textbf{BayesQuant Kruschke-like diagram}. Extends figure \ref{fig:daft_model} to give more detail into the model. Prior probability distributions are indicated with "distrograms", and the dependencies between them via arrows. The hyperparameters are represented by latin capital letters. Each $\beta$ will consist of a vector of dimensions $1 \times n$ where $n$ is the number of different factors in the category (i.e 2 for treatment, 6 for run, etc). The corresponding $x$ variable is a matrix of dimensions equal to $n \times n_peptides$ where the $i,j$ cell indicates if the $jth$ peptide is affected by the $ith$ factor (1) or not (0). Style based on \cite{Kruschke}.}
\label{fig:dbda}
\end{figure}

@@

\subsection{Sampling from the approximated posterior}
\label{subsec:basic_model}
